Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-04-30 overleaf/backend.tex

The taint analysis works on MPC Source, which lacks if-then-else control flow.
This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
%taint analysis with "taintedness" referring to the shared attribute.
Specifically, the compiler uses the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
\begin{enumerate}
\item Loop counters are always \texttt{plain}. \ben{I removed the reference to array bounds since those are not variables defined in MPC source}
\item If any variable on the right-hand side {\sf rhs} of an assignment is shared, then the assigned variable {\sf lhs} is \texttt{shared} following subtyping rule ${\sf rhs} <: {\sf lhs}$.
%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
\end{enumerate}

%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
The final rule is needed in examples like the following. In the below snippet {\sf sum!2} and {\sf sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
They are inferred as plaintext.
%cannot be marked as plaintext through simple taint analysis:
{\small
\begin{pythonn}
plaintext_array = [0, 1, 2, ...]
sum!1 = 0
for i in range(0, N):
    sum!2 = PHI(sum!1, sum!3)
    sum!3 = sum!2 + plaintext_array[i]
\end{pythonn}
}
%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}

@@

\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{True. To make matters worse, most of the explanation on this comes in the next section. But I'll leave as is for now but leave the comments too because it needs work.} \ben{Attempted to explain the naming, please review.}
\paragraph{Reshaping and raising dimensions:}
Raising the dimensions of a scalar or array uses the {\sf lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.  This function is also used for the scalar expansion of values which have been lifted out of FOR loops as described in \secref{sec:expansion}.  This function evaluates the expression for each permutation of indices along the dimensions and returns the resulting array in row-major order.  The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.  There is also a special case of the {\sf lift()} function which occurs when we are raising an array.  In this case, instead of concatenating the array for each index, \ana{can we remove "instead of concatenating the array"} \ben{I included it to show why the standard implementation of lift() doesn't work, if you don't think that's necessary then it can be removed.} we extend the array along all dimensions being raised which are not present in the array already.  For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: ${\sf A'}[n,m,d] = {\sf A}[n,m]$ for every $d$.  If the input array is already correctly sized it will be returned as-is.
%This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.  In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).
%\ana{Add the ref. to the figure. Also, try to rewrite. I understand what the paragraph is saying but it is too technical and does not convey the higher-level goal of these. Try first to state what the goal is, and then talk about the details such as the lambda funcs.} \ben{Rewrote the above paragraph to be less confusing/implementation-specific, please review.} \ana{Better, removing comment.}

>>>>

The taint analysis works on MPC Source, which lacks if-then-else control flow.
This significantly simplifies treatment as there is no need to handle conditionals and implicit flow.
%taint analysis with "taintedness" referring to the shared attribute.
Specifically, the compiler uses the following rules, which are standard in positive-negative qualifier systems (here \texttt{shared} is the positive qualifier and \texttt{plain} is the negative one):
\begin{enumerate}
\item Loop counters are always \texttt{plain}. %\ben{I removed the reference to array bounds since those are not variables defined in MPC source}\ana{Ok, resolved.}
\item If any variable on the right-hand side {\sf rhs} of an assignment is shared, then the assigned variable {\sf lhs} is \texttt{shared} following subtyping rule ${\sf rhs} <: {\sf lhs}$.
%\item If all variables on the right-hand side of an assignment are plaintext, then the assigned variable is plaintext \ana{Will be covered by the last rule.}
\item Any variables that cannot be determined as shared via the above rules are \texttt{plain}.
\end{enumerate}

%The first two rules are standard for taint analysis, and the third rule follows from the MPC problem statement. \ben{Is the explanation for the third rule correct?}  \ana{Yes, that is exactly it!}
%The final rule is needed in examples like the following. 
In the below snippet {\sf sum!2} and {\sf sum!3} form a dependency cycle and there is no \texttt{shared} value that flows to either one.
They are inferred as plaintext.
%cannot be marked as plaintext through simple taint analysis:
{\small
\begin{pythonn}
plaintext_array = [0, 1, 2, ...]
sum!1 = 0
for i in range(0, N):
    sum!2 = PHI(sum!1, sum!3)
    sum!3 = sum!2 + plaintext_array[i]
\end{pythonn}
}
%\ben{I think the above example is unnecessary and could be replaced by an explanation of how "untainted" variables are implicitly plaintext, but I couldn't think of a way to phrase that.}\ana{I left the example, we can leave it in the TR, remove from paper.}

@@

%\ben{The compiler and C++ internally uses ``lift'' as the terminology for raise\_dim(), but the paper does not :(...} \ana{True. To make matters worse, most of the explanation on this comes in the next section. But I'll leave as is for now but leave the comments too because it needs work.} \ben{Attempted to explain the naming, please review.} \ana{Marked as resolved.}
\paragraph{Reshaping and raising dimensions:}
Raising the dimensions of a scalar or array uses the {\sf lift()} utility function which takes a lambda for the raised expression and the dimensions of the output.  This function is also used for the scalar expansion of values which have been lifted out of FOR loops as described in \secref{sec:expansion}.  This function evaluates the expression for each permutation of indices along the dimensions and returns the resulting array in row-major order.  The lambda accepts an array of integers representing the index along each of the dimensions being raised, and the translation of the expression which is being raised replaces each of the dimension index variables with the relevant subscript of this array.  There is also a special case of the {\sf lift()} function which occurs when we are raising an array.  In this case, instead of concatenating the array for each index, %\ana{can we remove "instead of concatenating the array"} \ben{I included it to show why the standard implementation of lift() doesn't work, if you don't think that's necessary then it can be removed.} \ana{ok, left it in, marked as resolved.}
we extend the array along all dimensions being raised which are not present in the array already.  For example, when raising an array with dimensions $N \times M$ to an array with dimensions $N \times M \times D$, the input array will simply be extended along the $D$ dimension: ${\sf A'}[n,m,d] = {\sf A}[n,m]$ for every $d$.  If the input array is already correctly sized it will be returned as-is.
%This case only occurs when the dimensionality of the array is equal to the first dimensions of the array (any dimension which is being lifted to which is not in the array is after every dimension in the array) \ben{I don't know how to phrase the previous sentence more concisely}.  In this case, the array is simply extended along the new dimensions (if the array is already correctly sized it is returned as is).
%\ana{Add the ref. to the figure. Also, try to rewrite. I understand what the paragraph is saying but it is too technical and does not convey the higher-level goal of these. Try first to state what the goal is, and then talk about the details such as the lambda funcs.} \ben{Rewrote the above paragraph to be less confusing/implementation-specific, please review.} \ana{Better, removing comment.}
